---
title: "Multiple linear regression"
author: "Wing-Ho Ko"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
library(broom)
library(ggplot2)
```

In this demo we will investigate the relation between dissolved oxygen and physical factors such as temperature, salinity, and depth. The data we use come from a subset of the CalCOFI (California Cooperative Oceanic Fisheries Investigations) bottle database.

## Loading data

Load the subsetted data:

```{r}
CalCOFI <- read_csv("data/CalCOFI_sub.csv")
```

Inspect the dataset

```{r}
glimpse(CalCOFI)
```

## Simple visualizations

```{r}
ggplot(CalCOFI) + geom_point(aes(y=O_mmol_kg, x=depth_m), size=0.5)
```

```{r}
ggplot(CalCOFI) + geom_point(aes(y=O_mmol_kg, x=T_degC), size=0.5)
```

```{r}
ggplot(CalCOFI) + geom_point(aes(y=O_mmol_kg, x=salinity), size=0.5)
```

## Additive multiple linear regression

### Specifying and fitting the model

Let's try an additive model:

```{r}
O_lm_add <- lm(O_mmol_kg ~ depth_m + salinity + T_degC, CalCOFI)
```

### Extracting information

Extract the usual summary

```{r}
O_sm_add <- summary(O_lm_add)
print(O_sm_add)
```

Again, we can extract various pieces of information conveniently using `tidy()` and `glance()` from `broom`:

```{r}
O_lm_add_tbl <- tidy(O_lm_add)
O_lm_add_tbl
```

```{r}
O_lm_add_gl <- glance(O_lm_add)
O_lm_add_gl
```

Example: getting the value of r-squared programatically

```{r}
O_lm_add_gl$r.squared
```

### Making prediction

To make prediction, we first need to construct the data frame of new data. To do so manually:

```{r}
new_depth = seq(30, 40, 5)
new_temp = seq(10, 15, 2.5)
new_sal = seq(33, 34, 0.5)

phys_new <- data.frame(
  depth_m = rep(new_depth, times=length(new_temp) * length(new_sal)),
  T_degC = rep(new_temp, times=length(new_sal), each=length(new_depth)),
  salinity = rep(new_sal, each=length(new_depth) * length(new_temp))
)
```

```{r}
O_new <- predict(O_lm_add, phys_new)
O_pred <- cbind(phys_new, O_new)
O_pred
```

### Diagnostics

Diagnostics: residuals on y and fitted value on x:

```{r}
ggplot() + geom_point(aes(x=fitted(O_lm_add), y=resid(O_lm_add)), size=0.5)
```


## Interacting models

Consider adding interaction between salinity and temperature

```{r}
O_lm_TSint <- lm(O_mmol_kg ~ depth_m + salinity * T_degC, CalCOFI)
```

Print the summary:

```{r}
O_sm_TSint <- summary(O_lm_TSint)
print(O_sm_TSint)
```

Let's consider the same diagnostic plot

```{r}
ggplot() + geom_point(aes(x=fitted(O_lm_TSint), y=resid(O_lm_TSint)), size=0.5)
```

### Model selection

Since the two models are nested we can perform an F-test

```{r}
anova(O_lm_add, O_lm_TSint)
```

Thus it is probably a good idea to keep the interaction term. 

Alternatively, we can also investigate the AIC of the two models:

```{r}
AIC(O_lm_add, O_lm_TSint)
```

Again AIC suggests keeping the interaction term.

Finally, let's try to automatically select a model starting from the additive one: 

```{r}
O_lm_best <- step(O_lm_add, O_mmol_kg ~ depth_m * salinity * T_degC)
```

Let's try the "best" model that the stepwise procedure

```{r}
O_lm_best <- step(O_lm_add, O_mmol_kg ~ depth_m * salinity * T_degC)
```

Turns out the fully interacting model is selected. Printing out:

```{r}
summary(O_lm_best)
```
